Risks of Running Vulnerable Container Images for Data Analysis on HPC Clusters

### 2019-07-05 paper.tex
Repronim

\title{Use of Containers and Security threat to Clusters}

@@

\author[1]{Bhupinder Kaur}
\author[1]{Aiman Latif Hanna}
\author[1]{Tristan Glatard}

@@

\begin{abstract}

Today, containers, which provide operating system level virtualization, are very
popular. This is due to the need for fast development cycles, continuous deliv-
ery, and cost reduction of infrastructure. Moreover, demand for an isolated en-
vironment adds on to the popularity of containers. Indeed, containers accelerate
development cycles as compared to virtual machines and provide near-native performance.
Containers are isolated from each other but they share the host kernel. If the host
is compromised, all containers running on it are at risk. In this paper, we pre-
cisely study the security implications of containers. This includes the study of
vulnerabilities in container images, Linux kernel, and in the network. Here, we
first provide an introduction of containers, review of container’s security, and
later, we discuss how these security loopholes can be a risk to clusters.
\end{abstract}

\begin{keywords}
Containers; Docker; Singularity; Security; 
Clusters.
\end{keywords}

\section{Introduction}

Containers are very lightweight, flexible, and more resource-efficient than virtual
machines. They provide an operating system (OS) level virtualization, which
removes the overhead of having an extra OS layer. However, containers tightly
integrate with the host due to the sharing of the kernel, which raises security
concerns. Linux Containers (LXC) were introduced in 2008 and gained advantage 
over other Linux-based projects due to the integration of virtualization features 
into the upstream Linux kernel, which removed the need of applying patches to
Linux kernel and recompiling. LXC relies on cgroups and namespaces features
of the Linux kernel. Cgroups control the usage of system resources (memory, CPU,
network, etc) among a group of processes. Namespaces define what a process can
see by providing an isolated view of the operating system.

Later in 2013, Docker, which is an extension of LXC capabilities, was introduced.
Docker is a portable container engine used to pack an application
alongwith all its dependencies and libraries. The packed application is known as
a Docker image and it can be run on any Linux server. As Docker is also based
on LXC it also relies on cgroups for limiting resources and namespaces for
container isolation. All Docker containers are managed by the Docker engine
which is knows as Docker daemon. It is a service that runs on the host operating
system and requires root access to operate. Docker containers are
spawned as children of root-owned Docker daemon. Therefore,
users can gain escalated privileges by coercing Docker daemon,
which creates important security risks in shared environments. As a conse-
quence, Docker is usually not supported in multi-user environments.

Couple of years later, Singularity was introduced in 2016 with the main goal
of providing support for multi-user environment.
It offers mobility of compute by facilitating portable environments through 
a single image file, which makes Singularity containers different from 
Docker’s layered containers. Singularity takes snapshot, locks, and 
archives the developed application, which is known as Singularity image
Once Singularity image is created, it is hashed by using SHA256 hashing. 
Hence, it cannot be changed once it is created, consequently, it can be 
used for reproducing the results of scientific experiments.

Mobility of compute is defined as creating and maintaining a workflow on
a local machine that can be easily ported on other hosts, Linux distributions,
and/or cloud service providers. To achieve this, Singularity packs everything
that is required for the application to run in a distributable image. Then that
image can be copied, archived, and shared. Moreover, Singularity containers
are easily portable across different versions of the C library and different kernel
implementations.
Same features of Singularity, which are used for mobility of compute, facil-
itate reproducibility as well. Hash of the image is also stored along with the
image. All Singularity images can be stored at a central place is known as
Singularity hub and Docker images can be stored at Docker hub.

The remainder of this article describes security concerns of containers
and their chances of being exploited on clusters. It closes on a discussion
of ten simple rules that can be followed to release secure container images
for use in clusters.

\begin{figure}
\centering
        \includegraphics[width=\columnwidth]{Figures/virtualization.png}
        \caption{Architecture of Container-based
                Virtualization}
	\label{virtualization}
\end{figure}

\section{Container's Security Review}

In spite of the popularity of containers, there are various security concerns
as well. We categorize security of containers in two broad categories i.e container's 
internal security and security provided by the host which is known as
host hardening. In this article, we mainly focus on security concerns of 
Docker and Singularity containers as both are currently more popular than other 
OS-level virtualizations.

\subsection{Container's Internal Security}

We examine the internal security of containers based on the requirements given by
Reshetova et al.,2014 for comparing the security of a number of OS-level
virtualization techniques. According to this paper, an OS-level virtualization
technique security can be analyzed from its isolation from the host machine and from 
the technique used by it to limit resources. Apart from that, in this section, we discuss various security 
concerns added by vulnerabilities that are present in Docker images.

Isolation of containers from the host can be of different categorize such as process isolation, filesys-
tem isolation, device isolation, IPC isolation, network isolation. In this section, we discuss how
much justice is done to these isolations by Docker and Singularity containers.

\subsubsection{Process Isolation}

The main goal of process isolation is to prevent one container process from
seeing or interfering with another container process. In other words, it limits
9the permissions and visibility of a container process to processes running in other
containers. Docker achieves this goal through the use of namespaces in the Linux
kernel. For this purpose, Docker uses PID namespaces which isolate a process
with particular process ID from the host and other containers. Provided that,
it becomes difficult for an attacker to see processes running in other containers,
hence harder to attack them [24]. In contrast to the Docker, the PID namespace
is not isolated from the host by default in Singularity because it’s main goal is
to provide mobility of compute and reproducibility, not full isolation. But if
there is a need the user can separate the PID namespace using a command line
(or environment variable) setting.

\subsubsection{Filesystem Isolation}

Filesystem isolation is required to prevent illegitimate access to filesystems of
the host and containers. Docker uses filesystem namespaces, also called mount
namespaces, for achieving this goal. Filesystem namespaces hide filesystems
of the host and containers from other containers. However, some of the kernel
filesystems are not namespaced, so Docker containers mount them for operation,
for example, /sys, /proc/sys, /proc/sysrq - trigger, /proc/irq, and /proc/bus
[25]. This causes security concerns as Docker containers are directly able to
access host filesystems. Consequently, Docker provides two filesystem protection
mechanisms. First, Docker gives read-only permissions to containers for these
filesystems. Second, containers are not allowed to remount any filesystem within
containers. Additionally, Docker offers a copy-on-write filesystem mechanism,
which allows containers to write to their specific filesystems and changes are not
visible to other containers. On the other hand, files in Singularity containers
are not isolated using filesystem namespaces because a user inside a Singularity
container is the same user outside the container.

\subsubsection{Device Isolation}

Applications and kernel access devices through special files known as device
nodes. It is very crucial to limit the set of devices nodes that containers can
access. This is primarily because an attacker can own the whole system by
gaining access to some important device nodes such as /dev/mem, /dev/sd*,
or /dev/tty. Docker uses the Device Whitelist Controller [26] feature of con-
trol groups to limit the set of devices that a Docker container can see and use.
Additionally, Docker starts containers with nodev which prevents the use of
already created device nodes inside the image. Furthermore, Docker does not
allow containers to create new device nodes. However, some of the important
device nodes cannot be namespaced such as /dev/mem, /dev/sd* and kernel
modules. Direct access to these device nodes by containers possess serious se-
curity concerns. Besides that, if a Docker container is executed in privileged
mode then it gets access to all devices. Conversely, in Singularity devices are
also not isolated between the host and the container similar to filesystem due to
same user inside and outside the container. In Singularity by default the user
is allowed to interface all the devices of the host.
\subsubsection{IPC Isolation}

IPC is a set of objects through which processes communicate with each other
such as, shared memory segments, semaphores, message queues, and POSIX
message queues. IPC isolation is needed to prevent containers from accessing
or modifying data belonging to other containers, which is transmitted through
these objects. Docker utilizes IPC namespaces to assign an IPC namespace
to each container. Process in one IPC namespace cannot read or write IPC
resources of another IPC namespace.

\subsubsection{Network Isolation}

Isolation of containers network is very important to prevent network-based
attacks such as address resolution protocol (ARP) spoofing and Man-in-the-
Middle (MitM) Attack [27]. ARP spoofing is an attack, which associates the
attacker’s media access control (MAC) address with the internet protocol (IP)
address of another host. ARP is a stateless protocol and hosts save all ARP
replies even if they had not sent any ARP request for it [28], which becomes
a reason of attack. So, the host that has spoofed ARP response is not able to
verify whether it belongs to legitimate host or attacker and hence, it starts send-
ing packets at attacker’s MAC address. Docker uses network namespaces [29]
to isolate the network of Docker containers. Therefore, each Docker container
has its different IP address, IP routing tables, network devices, etc. Conse-
quently, Docker containers interact through the network interfaces with each
other as well as with the host [30]. On the contrary, all containers share the
same network bridge, which makes Docker vulnerable to ARP spoofing attack.
In addition to that, the Docker network bridge forwards all incoming packets
without any kind of filtering which makes it vulnerable to Mac flooding.
In Singularity, by default the container share network
with the host.

Network is also used to pull container images from the central repository.
Images are pulled in compressed format and then uncompressed on the host and
run into containers. This may include storing and processing potentially
untrusted code in images by container engine. The source code can
be tampered during transfer or at the source. If some part of the network
is compromised, an attacker can replace an image with malicious image, and
that image gets downloaded on the host. As the image is in a compressed
format, the attacker can cleverly craft the image (i.e. all zeros) and when it
gets decompressed on the host it can fill whole storage of host causing denial-
of-service (DoS) attack. Other possible attacks are code injection or replay
attacks. Additionally, the malicious image can be uploaded to the image
repository by an adversary. That image can be downloaded by millions of users infecting
millions of machines.

For mitigation of these attacks, Docker introduced content trust which al-
lowed signing images before pushing to Docker Hub. But this content trust can
be disabled and thus disabling image signature check. Another issue is with au-
tomated build and webhooks, which are described earlier, where compromised
GitHub account can lead to the execution of malicious code. According to the
experiment performed in [13] the malicious code was put in production within
5 minutes and 30 seconds of commit on GitHub. The Content trust provides
an environment where a single entity is trusted but in this case, trust is divided
among several external entities.
\subsubsection{Limiting of Resources}

A DoS attack occurs when intended users are not able to use the system or
network resources [32]. To launch DoS, the attacker floods targeted host or
network traffic with superfluous requests to overload the systems. Consequently,
the target crashes, hence disrupting normal execution of the system. To solve
this issue Docker uses cgroups. Cgroups restrict the amount of resources (CPU,
memory, and disk I/O) that are used by Docker containers, thus all containers
get their fair amount of resources.

\subsection{Image vulnerabilities}

There can be vulnerabilities present inside the image itself when it is downloaded
from image repository. According to the technical report [33] over 30\% official im-
ages have high-priority common vulnerabilities and exposures (CVE) identifiers
(IDs) and around 64\% have high or medium level CVE vulnerabilities. This
technical report also states that Docker images with the latest tag also have
vulnerabilities. These vulnerabilities are due to the outdated packages con-
tained by images which can be due to the use of an old base image or due to
12outdated code pulled during build. Vulnerabilities can also be present 
inside Singularity images as well. As these Docker images can also be
wrapped in Singularity images which automatically transfers vulnerabilities
between images.
Docker introduced Docker security scanning through which users can scan
images to check whether they contain vulnerability or not. However, this scan-
ning is limited to only private repositories and it is paid. The scan traverses
all layers of the image and then identifies packages and software in those layers.
Further, it checks vulnerabilities in these software components by taking their
Secure Hash Algorithms (SHAs) and then comparing against a standard list of
CVEs. This scan can take up to 1 to 24 hours depending on the image sizes.
Additionally, this scanning technique do not detect vulnerabilities which are not
mentioned in the standard CVE database.

\subsection{Host Hardening}

Out-of-the-box security provided by the Linux kernel to secure the host from
containers is known as host hardening. Linux provides Linux capabilities and
Linux Security Modules (LSM) to harden the security of the host system. Linux
capabilities divide the privileges of superuser into pieces and assign a subset of
these capabilities to specific processes. Whereas, LSM provides a framework for
Linux to support various security modules. Currently, three security modules
are officially integrated with Linux kernel which includes Security Enhancement
to the Linux system (SELinux), AppArmor, and Seccomp. Out of these three, 
only SELinux and AppArmor are supported by Docker. Docker integrates with 
Seccomp only if LXC are used.

\subsubsection{Linux Capabilities}

According to the main page of Linux capabilities, traditionally Unix systems
categorized processes as privileged processes and unprivileged processes. Priv-
ileged processes are root users with zero user id (UID), whereas unprivileged
users are normal users with nonzero UID. Privileged processes are exempted
from permission check whereas, unprivileged processes are liable to full per-
mission checks. Linux divides the privileges of superuser into different pieces,
known as capabilities, which can be independently enabled or disabled. As a
result, Docker can disable some of the capabilities of containers, thus improv-
ing the security of the system. As Docker containers share the kernel with the
host so, most of their tasks are done by the host. As a consequence, disabling
some of the capabilities in Docker containers do not affect their functionality.
For example, CAP\_NET\_ ADMIN capability allows configuration of the network
which can be disabled in Docker containers because all network configurations
are handled by Docker daemon. By default, most of the Linux capabilities are
disabled when Docker container is started, in order to secure the host system
from attackers.

\subsubsection{Linux Security Modules}

SELinux and AppArmor are both Media Access Control (MAC) solutions for 
enhancing security of Linux systems. MAC strongly separates all 
applications, which in turn decreases potential damage if an 
application is compromised. Unfortunately, Security provided by these modules 
is limited due to generic nature of profiles
provided by these security modules. For example, default SELinux profile
assigns same domain to all Docker containers which helps in protecting host from
containers but not containers from containers. Similarly, default Apparmor
profile provide full access of network system, file system and capabilities to
Docker containers. Solution can be writing specific profiles for the individual
containers.

@@

\subsection{Compute Canada Clusters}
Compute Canada is a national High Performance Computing (HPC) research platform.
It integrates with four regional partners namely ACENET, Calcul Québec, Compute
Ontario, and WestGrid for providing local support.
Compute Canada uses Slurm Workload Manager to schedule jobs.
Control groups are used to limit number of resources that can be used
by a particular user.
Compute Canada do not support Docker as its default configuration
runs containers as root. Whereas, Compute Canada supports Singularity
because it removes the drawback of Docker.
Compute Canada uses Lustre shared file system to provide high
bandwidth to its users.

@@

Following are the application frameworks that are using containers
on HPC clusters.

@@

Boutiques framework is used to publish, integrate, and execute third-party
applications across CBRAIN platform automatically. Boutiques framework is
employed by CONP. With Boutiques, the command-line of the application is
described through a flexible template, which is called descriptor. Descriptor de-
scribes the input that is required by the application and the output it produces.
It points to a Docker or Singularity image, where the application is installed.
Developers of the application produce Boutiques descriptor, and stores it along
with the application. Boutiques descriptor are described in JSON format. There
is a set of core tools which help in constructing, validating and executing bou-
tiques descriptors. During the application run-time, the execution platform
builds a command line by using input values given by the user and boutiques
descriptor of the application.

\subsubsection{Brain Imaging Data Structure (BIDS)}

 BIDS is a framework

\section{Containers on Clusters}

In this section, we discuss Canadian Open Neuroscience Platform (CONP) which is 
real use case of containers that involves running containers on clusters. 
CONP is a national platform
for sharing neuroscience research data and aims at creating an interactive inter-
face for neuroscientists and clinical neuroscience. The main goal of the CONP
platform is to store, process, and distribute massive amounts of data produced
by modern neuroscience. The effective data sharing by CONP increases data
accessibility and re-usability. Through CONP, neuroscientists remain updated
of on-going and past research efforts, which avoids unnecessary duplication.
CONP uses Compute Canada cluster for providing its services. So we take
this real use case to explain the effect of image vulnerabilities on clusters.

\sections{Cluster}

A cluster is an arrangement of multiple computers together in particular topol-
ogy to increase computational power, memory and reduce processing time than
personal computer. When a single computer becomes insufficient to use due
to the massive amount of data, cluster of computers can be used to run jobs
in parallel. Most importantly, clusters are the resources for high performing
computing (HPC). HPC refers to the use of clusters at large scale and parallel
processing techniques to solve large and difficult problems. Research, aca-
demics, and IT world share clusters for running jobs. Due to the wide use and
popularity of clusters, its security becomes a main concern. Below is the real
use case of clusters that deal with containers and might be at risk.

>>>>

\title{Risks of Running Vulnerable Container Images for Data Analysis on HPC
Clusters}

@@

\author[1]{Aiman Latif Hanna}
\author[1]{Bhupinder Kaur}
\author[1]{Tristan Glatard}

@@

@@

In this section we summarize a particular use case of containers which is Canadian
Open Neuroscience Platform (CONP). It is a national platform for sharing neuroscience
research data and aims at creating an interactive interface for neuroscientists and
clinical neuroscience. The main goal of the CONP platform is to store, process, and
distribute massive amounts of data produced by modern neuroscience.
Here, we summarize all the resources and software tools that are
used by CONP to function.

\subsection{Compute Canada Clusters}

Compute Canada is a national High Performance Computing (HPC) research platform.
It provides infrastructure and services for Advanced Research Computing (ARC).
Its goal is to bring excellence, ease, and innovation in research of Canadian
research institutions by providing storage, computing power, and required
softwares to carry out research.
It integrates with four regional partners namely ACENET, Calcul Québec, Compute
Ontario, and WestGrid for providing local support.
Compute Canada uses Slurm Workload Manager to schedule jobs.
Control groups are used to limit number of resources that can be used
by a particular user.
Compute Canada do not support Docker as its default configuration
runs containers as root. Whereas, Compute Canada supports Singularity
because it removes the drawback of Docker.
Compute Canada uses Lustre shared file system to provide high
bandwidth to its users.

@@

Following are the application frameworks in neuroscience that are using containers
on HPC clusters.

@@

Boutiques is a application sharing system. It is based on Linux containers
and can be used to publish, integrate, and execute third-party
applications across different platforms automatically. Boutiques framework is
employed by CONP. With Boutiques, the command-line of the application is
described through a flexible template, which is called descriptor. Descriptor
describes the input that is required by the application and the output it produces.
It points to a Docker or Singularity image, where the application is installed.
Developers of the application produce Boutiques descriptor, and stores it along
with the application. Boutiques descriptor are described in JSON format. There
is a set of core tools which help in constructing, validating and executing bou-
tiques descriptors. During the application run-time, the execution platform
builds a command line by using input values given by the user and boutiques
descriptor of the application.

\subsubsection{Brain Imaging Data Structure (BIDS) Apps}

BIDS Apps is a framework designed to share and execute neuroimaging
analysis pipelines. It improves usability, accesibility, and reproducibility
of different operating system users as well as HPC users.

\subsubsection{ReproNim}

ReproNim: A Center for Reproducible Neuroimaging Computation is a vision
to help neuroscientists in reproducing neuroimaging research.

\subsection{Container Images Used}

\subsection{Image Update Process}

If there exist vulnerabilities in the container image, then there are two ways
to proceed, either not to run that image at all, or run only after updating the
packages that have vulnerabilities. The latter option effects the reproducibility
of the image, which means that the image cannot be used to verify research
findings of other scientists. An alternative approach can be to trim unnecessary
packages from images, which in turn reduces the number of vulnerabilities
present in images.

\subsection{Scanning tools used}

There are various scanning tools that are available for scanning container
images like Vuls, Clair, ClairScanner, Anchore, Dagda, etc. Out of them we
took Anchore and Vuls for our experiments to scan container
images which are used by the above discussed application frameworks.

\subsubsection{Anchore}

Anchore is an end-to-end container security, and open-source compliance plat-
form. Anchore is a static scanning tool that analyses container images and
lists operating system packages,
unofficial packages, configuration files, and language modules [58]. In other
words, Anchore is a set of tools, which users can employ for scanning container
images and applying custom policies to them. It provides visibility and trans-
parency of the container environment. Anchore offers services in two ways:
Anchore Engine and Anchore Engine Command Line Interface (CLI).
Anchore Engine is available as a Docker image, which is available on Docker
Hub. Anchore Engine CLI provides command line interface in addition to An-
chore Engine REST API. Anchore Engine CLI can be installed directly on the
host and used to inspect images.

\subsubsection{Vuls}

Vuls is an open-source vulnerability scanner for Linux/FreeBSD, written in Go
language. It is an agentless vulnerability scanner and uses various
vulnerability databases listed in 4.4a. Using dynamic scanning technique, it offers
remote and local scan. Remote scan involves setting up only one machine, to
which other target servers are connected via SSH. Local scan involves running
the scan on the target server itself. Further, Vuls offer fast scan, fast root scan,
and deep scan.
Fast scan is performed without root privileges, whereas fast root scan is
performed with root privileges. The latter also uses checkrestart utility to de-
tect processes that are updated but not yet restarted because if they are not
restarted, updates are not effective. So fast root scan also detects these pro-
cesses. Both fast scan and fast root scan are performed offline without Internet
access, and put almost no load on the scan target server. Deep scan uses
root privileges for scanning, and is used to get a more detailed scan. It checks
changelogs, which puts more load on the target server. As shown in Figure 4.3
it is clear that vulnerability detection time of Vuls is much better than An-
chore. Vuls tool uses dynamic scanning, so the scanning time does not depend
on the image size.
Graph 4.5 shows percentage of vulnerabilities that are
covered by Vuls and Anchore tools. This graph can be explained with the help
of List 4.4, which shows vulnerability databases that are referred, by both Vuls
and Anchore. Anchore covers less percentage of vulnerabilities as compared
to Vuls because Anchore relies mainly on vulnerability databases published by
Linux distributions, whereas Vuls additionally refers to another sources such as
databases from Japan, US-CERT, etc.

\subsection{CBRAIN}

Canadian Brain Imaging Research Platform (CBRAIN) is a web portal that is used
by CONP to exploit services of Compute Canada. CBRAIN is a collaborative
neuroimaging research platform that is active in production since 2009. CBRAIN
provides transparent access to various distributed storage sites and computing
resources across the world. CBRAIN’s infrastructure consists of three main
layers, namely access layer, service layer, and infrastructure layer. The access
layer is for CBRAIN users, and can be accessed through any modern browser or
RESTful API [42]. The service layer consists of CBRAIN portal, which contains
metadata and databases. It provides information about users, their permissions,
resources etc. The infrastructure layer consists of data resources and computing
resources, which are connected through a network.
